\LARGE{\bf{Natural variation in teosinte at the domestication locus \emph{teosinte branched1} (\emph{tb1})}}\footnotemark

### 2014-05-23 manuscript.tex
incorporating edits

\documentclass[11pt]{article}
\usepackage{geometry} 
\usepackage{hyperref}
\usepackage[leftcaption]{sidecap}
\sidecaptionvpos{figure}{c} %position side caption
\usepackage{lineno}
\usepackage{lipsum}
\usepackage{indentfirst}
\usepackage{color}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage[pdftex]{graphicx}  
%\usepackage{caption}
\usepackage{natbib}
\usepackage{authblk}
\usepackage[table]{xcolor} %for table line color alteration
\usepackage{subscript}
\renewcommand{\baselinestretch}{1.5}
\bibliographystyle{AJB}
\renewcommand{\refname}{\centering{LITERATURE CITED}}

@@

\emph{Key results:} Genotyping results suggest the \emph{Hopscotch} element is at higher than expected frequency in teosinte. Analysis of linkage disequilibrium near \emph{tb1} does not support recent introgression of the \emph{Hopscotch} allele from maize into teosinte.  Population genetic signatures are consistent with selection on this locus revealing a potential ecological role for \emph{Hopscotch} in teosinte. Finally, two greenhouse experiments with teosinte do not suggest \emph{tb1} controls tillering in natural populations. \lev{do you think that we could change this at all to say something to make it sound less like hopscotch has absolutely no effect? I think it's likely that it was something with our experiments, or that it is the effect of other loci} \jri{other loci means tb1 doesn't solely control. if we have no faith in our experimental results, we shouldn't be writing the paper! it may be daylength, crowding whatever in our experiment, but as long as we don't think anything was wrong with data collection and analysis, I say keep as is.}

@@

\lev{does Hopscotch need to be italicized? Don't believe we had ever italicized it before.}

@@

In teosinte and other plants that grow at high population density, individuals detect competition from neighbors via the ratio of red to far-red light.  An increase in far-red relative to red light accompanies shading and triggers the shade avoidance syndrome: a suite of physiological and morphological changes such as reduced tillering, increased plant height and early flowering\lev{we saw a bunch of early flowering in phenotyping 2} \jri{were plants taller in pheno2?  were they more crowded?} \citep{KebromBrutnell2007}. The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} and other grasses and may therefore be crucial to the ecology of teosinte \citep{KebromBrutnell2007, LukensDoebley1999}.  In this study we aim to characterize the distribution of the \emph{Hopscotch} insertion in \emph{parviglumis}, \emph{mexicana}, and landrace maize, and to examine the phenotypic effects of the insertion in \emph{parviglumis}. We use a combination of PCR genotyping for the \emph{Hopscotch} element in our full panel and sequencing of two small regions upstream of \emph{tb1} in a subset of teosinte populations to explore patterns of genetic variation at this locus. Finally, we test for an association between the \emph{Hopscotch} element and tillering phenotypes in a population of \emph{parviglumis}.

@@

We examined evidence of introgression on chromosome 1 in these same four populations (EjuA, EjuB, MSA, SLO) using STRUCTURE \citep{Falush2003} and phased data from \citet{Pyhajarvi2013}, combined with the corresponding SNP data from a diverse panel of 282 maize lines \citep{Cook2012}. SNPs were anchored in a modified version of the IBM genetic map \citep{Gerke2013}. We created haplotype blocks using a custom Perl script \lev{is this where I should cite Matt's paper, or his code? Matt is was your 'chunkfy.pl script} \mbh{we should make this script available; my vote would be to get this submitted and during review get all scripts, etc... in GitHub or some other public repository} \jri{i'm ok with wiating to do after submit; I'm not OK with publication until everything is on github} that grouped SNPs separated by less than 5kb into haplotypes. We ran STRUCTURE at K=2 under the linkage model, performing 3 replicates with an MCMC burn-in of 10,000 steps and 50,000 steps post burn-in.

@@

% Figure 3 STRUCTURE Plot 4 Pops
%-------------------------------------------------------------------
\begin{figure*}[!t]
  \begin{center}
   \includegraphics[width=90mm]{Fig3Structure.pdf}
    \caption{STRUCTURE assignment to maize across a section of chromosome 1. The dotted lines mark the beginning of the sequenced region 50kb upstream (Sequenced region 2) and the end of the \emph{tb1} ORF. } \jri{ we use 50kb sometimes and 66kb elsehwere. is this typo, or did we really only consider 50kb? }
\label{Fig3Structure}
  \end{center}
\end{figure*}
%-------------------------------------------------------------------

@@

In maize, alleles at domestication loci (\emph{RAMOSA1}, \citealt{SigmonVollbrecht2010}; \emph{barren stalk1}, \citealt{Gallavotti2004}; and \emph{grassy tillers1}, \citealt{Whipple2011}) are thought to have been selected from standing variation, suggesting that diversity already present in teosinte may have played an important role in maize domestication. The \emph{teosinte branched1} gene is one of the best characterized domestication loci, and, while previous studies have suggested that differences in plant architecture between maize and teosinte are a result of selection on standing variation at this locus \citep{Clark2006, Studer2011}, little is known about natural variation at this locus and its ecological role in teosinte. \jri{can we say this with the weber paper and studer allelic series papers? perhaps rephrase?}

@@

While gene flow from crops into their wild relatives is well-known, \citep{Ellstrand1999, Zhang2009, Thurber2010, Baack2008, Hubner2012, Wilkes1977, VanHeerwaarden2011, Barrett1983}, our results are more consistent with \citet{Hufford2013} who found resistance to introgression from maize into teosinte around domestication loci \citet{Hufford2013}. 
We find no evidence of recent introgression in our analyses. Clustering in our NJ trees do not reflect the pattern expected if maize alleles at the \emph{tb1} locus had introgressed into populations of teosinte.  Moreover, there is no signature of elevated LD in the \emph{tb1} region relative to the rest of chromosome 1, and Bayesian assignment to a maize cluster in this region is both low and below the chromosome-wide average (Fig. \ref{Fig3Structure}, Table \ref{Table4Q}). Together, these data point to an explanation other than recent introgression for the high observed frequency of \emph{Hopscotch} in a subset of our \emph{parviglumis} populations. \jri{ not worth includind in ms. but interesting the structure plot shows an uptick of introgression right around the hopscotch }

@@

% Figure S1. Cartoon representation of tb1 locus
%-------------------------------------------------------------------
\begin{figure*}[!t]
  \begin{center}
   \includegraphics[width=140mm]{FigS1LocusCartoon.pdf}
   \caption{Representation of the upstream regulatory region of \emph{tb1}, showing the \emph{tb1} coding region (green) and the \emph{Hopscotch} insertion (red). Arrows show the location of primer sets; in black, primers used for amplification and sequencing (Sequenced region 1; within the 5' UTR, and sequenced region 2; 66,169 bp upstream from the tb1 ORF); in blue, primers used to genotype the \emph{Hopscotch} insertion.}
\label{FigS1Locus}
  \end{center}
\end{figure*}
%-------------------------------------------------------------------

>>>>

\documentclass[11pt]{article}
\usepackage{geometry} 
\usepackage{hyperref}
\usepackage[leftcaption]{sidecap}
\sidecaptionvpos{figure}{c} %position side caption
\usepackage{lineno}
\usepackage{lipsum}
\usepackage{indentfirst}
\usepackage{color}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage[pdftex]{graphicx}  
\usepackage{caption}
\usepackage{natbib}
\usepackage{authblk}
\usepackage[table]{xcolor} %for table line color alteration
\usepackage{subscript}
\renewcommand{\baselinestretch}{1.5}
\bibliographystyle{AJB}
\renewcommand{\refname}{\centering{LITERATURE CITED}}

@@

\emph{Key results:} Genotyping results suggest the \emph{Hopscotch} element is at higher than expected frequency in teosinte. Analysis of linkage disequilibrium near \emph{tb1} does not support recent introgression of the \emph{Hopscotch} allele from maize into teosinte.  Population genetic signatures are consistent with selection on this locus revealing a potential ecological role for \emph{Hopscotch} in teosinte. Finally, two greenhouse experiments with teosinte do not suggest \emph{tb1} controls tillering in natural populations.
\vspace*{0.2cm}

@@

@@

In teosinte and other plants that grow at high population density, individuals detect competition from neighbors via the ratio of red to far-red light.  An increase in far-red relative to red light accompanies shading and triggers the shade avoidance syndrome: a suite of physiological and morphological changes such as reduced tillering, increased plant height and early flowering \citep{KebromBrutnell2007}. The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} and other grasses and may therefore be crucial to the ecology of teosinte \citep{KebromBrutnell2007, LukensDoebley1999}.  In this study we aim to characterize the distribution of the \emph{Hopscotch} insertion in \emph{parviglumis}, \emph{mexicana}, and landrace maize, and to examine the phenotypic effects of the insertion in \emph{parviglumis}. We use a combination of PCR genotyping for the \emph{Hopscotch} element in our full panel and sequencing of two small regions upstream of \emph{tb1} in a subset of teosinte populations to explore patterns of genetic variation at this locus. Finally, we test for an association between the \emph{Hopscotch} element and tillering phenotypes in a population of \emph{parviglumis}.

@@

We examined evidence of introgression on chromosome 1 in these same four populations (EjuA, EjuB, MSA, SLO) using STRUCTURE \citep{Falush2003} and phased data from \citet{Pyhajarvi2013}, combined with the corresponding SNP data from a diverse panel of 282 maize lines \citep{Cook2012}. SNPs were anchored in a modified version of the IBM genetic map \citep{Gerke2013}. We created haplotype blocks using a custom Perl script that grouped SNPs separated by less than 5kb into haplotypes. We ran STRUCTURE at K=2 under the linkage model, performing 3 replicates with an MCMC burn-in of 10,000 steps and 50,000 steps post burn-in.

@@

% Figure 3 STRUCTURE Plot 4 Pops
%-------------------------------------------------------------------
\begin{figure*}[!t]
  \begin{center}
   \includegraphics[width=90mm]{Fig3Structure.pdf}
    \caption{STRUCTURE assignment to maize across a section of chromosome 1. The dotted lines mark the beginning of the sequenced region 66kb upstream (Region 2) and the end of the \emph{tb1} ORF. }
\label{Fig3Structure}
  \end{center}
\end{figure*}
%-------------------------------------------------------------------

@@

In maize, alleles at domestication loci (\emph{RAMOSA1}, \citealt{SigmonVollbrecht2010}; \emph{barren stalk1}, \citealt{Gallavotti2004}; and \emph{grassy tillers1}, \citealt{Whipple2011}) are thought to have been selected from standing variation, suggesting that diversity already present in teosinte may have played an important role in maize domestication. The \emph{teosinte branched1} gene is one of the best characterized domestication loci, and, while previous studies have suggested that differences in plant architecture between maize and teosinte are a result of selection on standing variation at this locus \citep{Clark2006, Studer2011}, much remains to be discovered regarding natural variation at this locus and its ecological role in teosinte.

@@

While gene flow from crops into their wild relatives is well-known, \citep{Ellstrand1999, Zhang2009, Thurber2010, Baack2008, Hubner2012, Wilkes1977, VanHeerwaarden2011, Barrett1983}, our results are more consistent with \citet{Hufford2013} who found resistance to introgression from maize into teosinte around domestication loci. 
We find no evidence of recent introgression in our analyses. Clustering in our NJ trees do not reflect the pattern expected if maize alleles at the \emph{tb1} locus had introgressed into populations of teosinte.  Moreover, there is no signature of elevated LD in the \emph{tb1} region relative to the rest of chromosome 1, and Bayesian assignment to a maize cluster in this region is both low and below the chromosome-wide average (Fig. \ref{Fig3Structure}, Table \ref{Table4Q}). Together, these data point to an explanation other than recent introgression for the high observed frequency of \emph{Hopscotch} in a subset of our \emph{parviglumis} populations.

@@

% Figure S1. Cartoon representation of tb1 locus
%-------------------------------------------------------------------
\begin{figure*}[!t]
  \begin{center}
   \includegraphics[width=140mm]{FigS1LocusCartoon.pdf}
  %\caption{Representation of the upstream regulatory region of \emph{tb1}, showing the \emph{tb1} coding region (green) and the \emph{Hopscotch} insertion (red). Arrows show the location of primer sets; in black, primers used for amplification and sequencing (Sequenced region 1; within the 5' UTR, and sequenced region 2; 66,169 bp upstream from the tb1 ORF); in blue, primers used to genotype the \emph{Hopscotch} insertion.}
\label{FigS1Locus}
  \end{center}
\end{figure*}
%-------------------------------------------------------------------
